Maximum Entropy Spectral Analysis: a case study

### 2021-03-23 paper/mesa.tex
Eqrefs

without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows. The latter definition in (\ref{eq:psd-f-definition}) gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.  

@@

%In this context, the role played by evidence should be clear: it dictates the probability assignment in order to make the result `more certain' or `less informative'. Any assignment different from the `most uncertain' is violating the principle of it to being non committal to non-available evidence. We can only assign a more certain probability, i.e. less informative, if we have evidence for it. If we have none, the result has to be the most unpredictable, i.e. the most informative. From the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it.
\par
MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty using evidence as constraint. 
This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized. $H$ has to be a `generalization' of information defined in equation (\ref{eq:information}) to be something like the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
%
%Some properties for $H$ can be worked out easily. 
%From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$.
%Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes.
%Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of "getting it wrong" increases as $\frac{N -1}{N}$. So when equal probabilities are assigned, we want
%\begin{equation}
%    \nonumber
%    H\left[\frac{1}{N}, \dots, \frac{1}{N}\right]
%\end{equation}
%to increase monotonically with N.
%
%We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
%Last is composition property. Consider three events with probability $p_1, p_2, p_3$ such that $p_2 + p_3 = 1 - p_1 = q$. We require 
%\begin{equation}
%    H\left(p_1, p_2, p_3\right) = H(p_1, q) + qH\left(\frac{p_2}{q}, \frac{p_3}{q}\right),
%\end{equation}
%so that uncertainty is additive.

@@

%This interpretation is straightforward: if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'. The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum. Such an assignment is in agreement with our hypothesis of being `non committal' with unavailable evidence, since \ref{eq:entropy} is all we need to assign probabilities uniquely. Evidence can be added in form of constraints in the maximization.

@@

%First of all, we want to show that, in case in which our only constraint is exhaustiveness
%$\sum_{i = 1}^{N}p_i = 1 $,
%MAXENT reduces to indifference principle as it should be. The variational problem is defined as
%\begin{equation}\label{eq:entropyNormConst}
%    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
%\end{equation}
%with
%\begin{equation}
%    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
%\end{equation}
%so that equation \ref{eq:entropyNormConst} reduce to
%\begin{equation}\nonumber 
%    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
%\end{equation}
%The solution of the previous equation is easily worked out requiring every term to be zero: 
%\begin{equation}\nonumber 
%    p_i = e^{-\lambda - 1}.
%\end{equation}
%Applying the normalization condition, it is immediate to show that
%\begin{equation}
%    p_i = \frac{1}{N}.
%\end{equation}
%MAXENT reduces to the IP when only the normalization constraint is used.
%
%In the second example, we will add the constraint that the distribution has a known mean value: $\int \d{x} x p(x) = \bar{x}$. For simplicity we will work in the continuous case.
%The variational problem can be rewritten as: 
%\begin{equation}
%    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
%\end{equation}
%with constraints: 
%\begin{equation*}
%    \int \d{x} \; p(x)  = 1 \text{  and  }
%    \int \d{x} \; x p(x) = \bar x.
%\end{equation*}
%For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval.
%
%The general solution for $p(x)$ is:
%\begin{equation}
%    p(x) = e^{-1 - \lambda - \mu x} 
%\end{equation}
%where $\lambda$ and $\mu$ are two Lagrange multipliers
%The normalization conditions implies 
%\begin{equation}\nonumber 
%   1 =  \int \d{x} \; p(x) = e^{-1-\lambda}\int e^{-\mu x} dx, 
%\end{equation}
%so that: 
%\begin{equation}\nonumber 
%    e^{-1 -\lambda} = \mu.
%\end{equation}
%The fixed average condition provides a value for $\mu$ in terms of $\bar{x}$:
%\begin{equation}
%	\mu = \frac{1}{\bar{x}}
%\end{equation}
%The solution is then an exponential distribution:
%\begin{equation}
%    p(x) = \frac{1}{\bar{x}} \exp\left(-\frac{x}{\bar{x}}\right). \nonumber 
%\end{equation}

@@

In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
would reduce to a single Fourier transform.
However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation
is impossible.
Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
}, so that only a few values for the autocorrelation function can actually be computed reliably.
This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function? MAXENT can guide use in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method. The latter assumes that the unknown values of the autocorrelation are $0$. Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.
\par
As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~(\ref{eq:entropy}), is maximized 
subject to some problem-specific constraints. 
In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).
\par
Before doing so, there is a technical problem to solve: the definition of entropy (Eq.~\ref{eq:entropy}) depends on a probability distribution, 
not on the PSD. The variational problem can be formulated in terms of the power spectral density $S(f)$ alone by
considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA} \sschmidt{Which form has this filter?}. 
The entropy gain obtained by filtering white noise with a sampling rate $\Delta t$ and Nyquist frequency $Ny \equiv \frac{1}{2 \Delta t}$ is:
\begin{equation}\label{eq:EntropyGain}
    \Delta H = \int_{- Ny}^{Ny}\log S(f) df\,.
\end{equation}

@@

Thus, by maximizing eq.~\ref{eq:EntropyGain} with constraints in eq.~\ref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
This approach on PSD computation provide a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
Most importantly, the variational problem admits a closed-form analytical expression for $S(f)$.
The expression was first found by Burg~\cite{burg1975maximum}:
\begin{equation}\label{eq:MESApsd}
    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
\end{equation}
where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1.
The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
The coefficients $P_N$ and $a_s (s > 0)$ are to be determined by an iterative process (called Burg's algorithm). The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:MESA_solution}.

@@

To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}. This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~(\ref{eq:AR_p}): 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
\end{align}
Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed a wide-sense stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S_{AR(p)}(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
Eq.~(\ref{eq:ARspectrum}) and Eq.~(\ref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
process that model the observed process.

@@

\subsection{Forecasting} \label{sec:forecasting}
The link between MESA and autoregressive processes is of particular interest. Given the solution to the Levinson recursion to determine the $a_k$, see Appendix A, we have automatically the coefficients of the AR with equivalent PSD, hence we are able to exploit Eq.~(\ref{eq:AR_p}) to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
Indeed, an AR($p$) process models the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation:
\begin{align}\label{eq:p_forecast}
	p&(x_t|x_{t-1}, \ldots , x_{t-p}) \nonumber\\
	&= \frac{1}{\sigma\sqrt{2\pi}} \exp\left[-\frac{1}{2} \left(\frac{x_t - \sum_{i = 1}^p b_i x_{t-i}}{\sigma}\right)^2\right].
\end{align}
The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computation.

@@

\item \textbf{Optimum Bayes Decision rule} 
The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
time series for the process that has to be described as an autoregressive process of order $m$ (AR(m)) with m to be determined. The OBD is obtained chosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$. Each
hypothesis is uniquely determined from both the length and the values of the filter under study. Chosing a Likelihood of the form \ref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e. maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of \wdp{please check this equation}:

@@

Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson \cite{doi:10.1002/sapm1946251261} recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached. Then, the order $m$ is selected to be the one that minimizes the loss function.

@@

For each loss function, we compute the ensemble-averaged PSD (as well as the $90\%$ confidence level) and the ensemble-averaged frequency dependence of the error:
\begin{equation}\label{eq:ensemble_averaged_error}
	r(f) = \sum_i \frac{|S_i(f) - S(f)|}{S(f)} \;.
\end{equation}
Figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD for the three loss functions, 
as well as the ensemble-averaged error from Eq.~(\ref{eq:ensemble_averaged_error}).
Furthermore, Fig.~\ref{fig:optcomparison} displays the correlation between $r_i$ and the autoregressive 
order $m_i$ chosen for each independently drawn time series.

@@

\begin{figure}
	\centering
	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
	\caption{Relative error $r$ (as in eq.~(\ref{eq:freq_error})) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with gaussian PSD given in Fig.~\ref{fig:autocorr}. Different colours refers to different choices for the loss function: in red $\mathcal{L}_{\rm FPE}(m)$, in green $\mathcal{L}_{\rm OBD}(m)$ and in black $\mathcal{L}_{\rm CAT}(m)$. The top and right histograms show the marginal distributions.}
	\label{fig:optcomparison}
\end{figure}

@@

Therefore, the choice of loss function, at least in between CAT and FPE, depends on the problem one  is attempting to solve.
\begin{figure}[t]
        \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates_peaks.pdf}
        \caption{Details of peaks of the spectrum and their reconstruction with every optimizer}
        \label{fig:ligoPeaks}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/error_length_contour.pdf}
    \caption{For each of the $500$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\ref{eq:freq_error}) against the length $m$ of autoregressive process. The time series are randomly drawn with a the analytical LIGO PSD in fig.`\ref{fig:ligospectrum}. Different colors refers to different choices for the loss function. Histograms for the distribution od the individual quantities are also represented.}
    \label{fig:LigoOrderError}
\end{figure}

@@

\pagebreak
\appendix
\section{MESA solution} \label{sec:MESA_solution}
\subsection{MESA solution}
We derive the expression for the MAXENT spectral estimator following the approach proposed by Burg \cite{burg1975maximum}.
Unlike the standard approach, we do not enforce the constraints in eq.~\ref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
We write instead the PSD $S(f)$ as the Fourier Transform of the sample autocorrelation function: 
\begin{equation}
    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t},
\end{equation}
and the entropy gain expression eq.~\ref{eq:EntropyGain} becomes:
\begin{equation}
    \Delta H = \int_{-Ny}^{Ny}  
    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t} 
    \right) df.
\end{equation}
Note that this expression already takes into account the contraints in eq.~\ref{eq:MaxConstraint}.

@@

Some properties for the coefficients can be worked out easily. First, since $S(f)$ is positive definite the $\lambda_s$ show the property 
\begin{equation}
	\nonumber 
	\lambda_s = \lambda_{-s}^*. 
\end{equation}
The second property is obtained considering that the autocorrelation function $r_n$ can only be computed for a finite time interval $n \in [-N, N]$ and that the PSD estimation must not depend on the unavailable values $r_n$: this is part of the constraint in eq.~\ref{eq:MaxConstraint}
This requirement can be implemented as:  
\begin{equation}\nonumber 
    \frac{\delta H}{\delta \bar r_s} = 0 \text{ for } \vert s \vert > N,
\end{equation}
that means 
\begin{equation}
\nonumber 
\lambda_s = 0 \text{ for } \vert s \vert > N. 
\end{equation}

>>>>

without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows. The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.  

@@

%In this context, the role played by evidence should be clear: it dictates the probability assignment in order to make the result `more certain' or `less informative'. Any assignment different from the `most uncertain' is violating the principle of it to being non committal to non-available evidence. We can only assign a more certain probability, i.e. less informative, if we have evidence for it. If we have none, the result has to be the most unpredictable, i.e. the most informative. From the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it.
\par
MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty using evidence as constraint. 
This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized. $H$ has to be a `generalization' of information defined in equation \eqref{eq:information} to be something like the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
%
%Some properties for $H$ can be worked out easily. 
%From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$.
%Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes.
%Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of "getting it wrong" increases as $\frac{N -1}{N}$. So when equal probabilities are assigned, we want
%\begin{equation}
%    \nonumber
%    H\left[\frac{1}{N}, \dots, \frac{1}{N}\right]
%\end{equation}
%to increase monotonically with N.
%
%We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
%Last is composition property. Consider three events with probability $p_1, p_2, p_3$ such that $p_2 + p_3 = 1 - p_1 = q$. We require 
%\begin{equation}
%    H\left(p_1, p_2, p_3\right) = H(p_1, q) + qH\left(\frac{p_2}{q}, \frac{p_3}{q}\right),
%\end{equation}
%so that uncertainty is additive.

@@

%This interpretation is straightforward: if we compare it with Eq.~(\eqref{eq:information}), we see that the entropy is defined as the 'average information'. The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum. Such an assignment is in agreement with our hypothesis of being `non committal' with unavailable evidence, since \eqref{eq:entropy} is all we need to assign probabilities uniquely. Evidence can be added in form of constraints in the maximization.

@@

%First of all, we want to show that, in case in which our only constraint is exhaustiveness
%$\sum_{i = 1}^{N}p_i = 1 $,
%MAXENT reduces to indifference principle as it should be. The variational problem is defined as
%\begin{equation}\label{eq:entropyNormConst}
%    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
%\end{equation}
%with
%\begin{equation}
%    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
%\end{equation}
%so that equation \eqref{eq:entropyNormConst} reduce to
%\begin{equation}\nonumber 
%    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
%\end{equation}
%The solution of the previous equation is easily worked out requiring every term to be zero: 
%\begin{equation}\nonumber 
%    p_i = e^{-\lambda - 1}.
%\end{equation}
%Applying the normalization condition, it is immediate to show that
%\begin{equation}
%    p_i = \frac{1}{N}.
%\end{equation}
%MAXENT reduces to the IP when only the normalization constraint is used.
%
%In the second example, we will add the constraint that the distribution has a known mean value: $\int \d{x} x p(x) = \bar{x}$. For simplicity we will work in the continuous case.
%The variational problem can be rewritten as: 
%\begin{equation}
%    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
%\end{equation}
%with constraints: 
%\begin{equation*}
%    \int \d{x} \; p(x)  = 1 \text{  and  }
%    \int \d{x} \; x p(x) = \bar x.
%\end{equation*}
%For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval.
%
%The general solution for $p(x)$ is:
%\begin{equation}
%    p(x) = e^{-1 - \lambda - \mu x} 
%\end{equation}
%where $\lambda$ and $\mu$ are two Lagrange multipliers
%The normalization conditions implies 
%\begin{equation}\nonumber 
%   1 =  \int \d{x} \; p(x) = e^{-1-\lambda}\int e^{-\mu x} dx, 
%\end{equation}
%so that: 
%\begin{equation}\nonumber 
%    e^{-1 -\lambda} = \mu.
%\end{equation}
%The fixed average condition provides a value for $\mu$ in terms of $\bar{x}$:
%\begin{equation}
%	\mu = \frac{1}{\bar{x}}
%\end{equation}
%The solution is then an exponential distribution:
%\begin{equation}
%    p(x) = \frac{1}{\bar{x}} \exp\left(-\frac{x}{\bar{x}}\right). \nonumber 
%\end{equation}

@@

In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
would reduce to a single Fourier transform.
However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation
is impossible.
Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
}, so that only a few values for the autocorrelation function can actually be computed reliably.
This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function? MAXENT can guide use in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method. The latter assumes that the unknown values of the autocorrelation are $0$. Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.
\par
As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~\eqref{eq:entropy}, is maximized 
subject to some problem-specific constraints. 
In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).
\par
Before doing so, there is a technical problem to solve: the definition of entropy (Eq.~\ref{eq:entropy}) depends on a probability distribution, 
not on the PSD. The variational problem can be formulated in terms of the power spectral density $S(f)$ alone by
considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA} \sschmidt{Which form has this filter?}. 
The entropy gain obtained by filtering white noise with a sampling rate $\Delta t$ and Nyquist frequency $Ny \equiv \frac{1}{2 \Delta t}$ is:
\begin{equation}\label{eq:EntropyGain}
    \Delta H = \int_{- Ny}^{Ny}\log S(f) df\,.
\end{equation}

@@

Thus, by maximizing eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
This approach on PSD computation provide a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
Most importantly, the variational problem admits a closed-form analytical expression for $S(f)$.
The expression was first found by Burg~\cite{burg1975maximum}:
\begin{equation}\label{eq:MESApsd}
    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
\end{equation}
where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1.
The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
The coefficients $P_N$ and $a_s (s > 0)$ are to be determined by an iterative process (called Burg's algorithm). The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:MESA_solution}.

@@

To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\eqref{eq:MESApsd}. This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~(\eqref{eq:AR_p}): 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
\end{align}
Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed a wide-sense stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S_{AR(p)}(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
Eq.~(\eqref{eq:ARspectrum}) and Eq.~(\eqref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
process that model the observed process.

@@

\subsection{Forecasting} \label{sec:forecasting}
The link between MESA and autoregressive processes is of particular interest. Given the solution to the Levinson recursion to determine the $a_k$, see Appendix A, we have automatically the coefficients of the AR with equivalent PSD, hence we are able to exploit Eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
Indeed, an AR($p$) process models the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation:
\begin{align}\label{eq:p_forecast}
	p&(x_t|x_{t-1}, \ldots , x_{t-p}) \nonumber\\
	&= \frac{1}{\sigma\sqrt{2\pi}} \exp\left[-\frac{1}{2} \left(\frac{x_t - \sum_{i = 1}^p b_i x_{t-i}}{\sigma}\right)^2\right].
\end{align}
The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computation.

@@

\item \textbf{Optimum Bayes Decision rule} 
The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
time series for the process that has to be described as an autoregressive process of order $m$ (AR(m)) with m to be determined. The OBD is obtained chosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$. Each
hypothesis is uniquely determined from both the length and the values of the filter under study. Chosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e. maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of \wdp{please check this equation}:

@@

Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson \cite{doi:10.1002/sapm1946251261} recursion until $M_{max}$, as given in Eq.~\eqref{eq:MMAx}, iterations are reached. Then, the order $m$ is selected to be the one that minimizes the loss function.

@@

For each loss function, we compute the ensemble-averaged PSD (as well as the $90\%$ confidence level) and the ensemble-averaged frequency dependence of the error:
\begin{equation}\label{eq:ensemble_averaged_error}
	r(f) = \sum_i \frac{|S_i(f) - S(f)|}{S(f)} \;.
\end{equation}
Figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD for the three loss functions, 
as well as the ensemble-averaged error from Eq.~\eqref{eq:ensemble_averaged_error}.
Furthermore, Fig.~\ref{fig:optcomparison} displays the correlation between $r_i$ and the autoregressive 
order $m_i$ chosen for each independently drawn time series.

@@

\begin{figure}
	\centering
	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
	\caption{Relative error $r$ (as in eq.~\eqref{eq:freq_error}) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with gaussian PSD given in Fig.~\ref{fig:autocorr}. Different colours refers to different choices for the loss function: in red $\mathcal{L}_{\rm FPE}(m)$, in green $\mathcal{L}_{\rm OBD}(m)$ and in black $\mathcal{L}_{\rm CAT}(m)$. The top and right histograms show the marginal distributions.}
	\label{fig:optcomparison}
\end{figure}

@@

Therefore, the choice of loss function, at least in between CAT and FPE, depends on the problem one  is attempting to solve.
\begin{figure}[t]
        \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates_peaks.pdf}
        \caption{Details of peaks of the spectrum and their reconstruction with every optimizer}
        \label{fig:ligoPeaks}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/error_length_contour.pdf}
    \caption{For each of the $500$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\eqref{eq:freq_error}) against the length $m$ of autoregressive process. The time series are randomly drawn with a the analytical LIGO PSD in fig.`\ref{fig:ligospectrum}. Different colors refers to different choices for the loss function. Histograms for the distribution od the individual quantities are also represented.}
    \label{fig:LigoOrderError}
\end{figure}

@@

\pagebreak
\appendix
\section{Computing the PSD} 
\subsection{MESA solution}\label{sec:MESA_solution}
We derive the expression for the MAXENT spectral estimator following the approach proposed by Burg \cite{burg1975maximum}.
Unlike the standard approach, we do not enforce the constraints in eq.~\eqref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
We write instead the PSD $S(f)$ as the Fourier Transform of the sample autocorrelation function: 
\begin{equation}
    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t},
\end{equation}
and the entropy gain expression eq.~\eqref{eq:EntropyGain} becomes:
\begin{equation}
    \Delta H = \int_{-Ny}^{Ny}  
    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t} 
    \right) df.
\end{equation}
Note that this expression already takes into account the contraints in eq.~\eqref{eq:MaxConstraint}.

@@

Some properties for the coefficients can be worked out easily. First, since $S(f)$ is positive definite the $\lambda_s$ show the property 
\begin{equation}
	\nonumber 
	\lambda_s = \lambda_{-s}^*. 
\end{equation}
The second property is obtained considering that the autocorrelation function $r_n$ can only be computed for a finite time interval $n \in [-N, N]$ and that the PSD estimation must not depend on the unavailable values $r_n$: this is part of the constraint in eq.~\eqref{eq:MaxConstraint}
This requirement can be implemented as:  
\begin{equation}\nonumber 
    \frac{\delta H}{\delta \bar r_s} = 0 \text{ for } \vert s \vert > N,
\end{equation}
that means 
\begin{equation}
\nonumber 
\lambda_s = 0 \text{ for } \vert s \vert > N. 
\end{equation}
